---
title: "Random Forest"
description: |
 Predicting Bankrupcty using Random Forest"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# [^4]  Company Bankruptcy Prediction Dataset

[^4]: <https://www.kaggle.com/fedesoriano/company-bankruptcy-prediction>

<center>
# What is Random Forest?

#### Random Forest is a algorithm that comes to a conclusion using many different decisions trees. The most common outcome for each indiviual decision tree helps to create the final output. While this algorithm may seem complex the core of this algorithm decision trees, have been used my most wether knowingly or not.

</center>

```{r}
library(tidyverse)

library(DBI)
 bankdata <- read.csv("C:/Users/Jesse/Desktop/JFlowers R/MyWebsite/data/bankdata.csv")
```


```{r,echo=TRUE}
library(caret)
trainIndex <- createDataPartition(bankdata$Bankrupt., p = .6, list = FALSE, times = 1)
 
```


## Grab The Data
```{r,echo=TRUE}
#grab the data
BankTrain <- bankdata[ trainIndex,]
BankTest  <- bankdata[-trainIndex,]
```



## Perform PreProcessing on Data
```{r,echo=TRUE,results='hide',message=FALSE}
preProcValues <- preProcess(BankTrain, method = c("center", "scale"))

trainTransformed <- predict(preProcValues, BankTrain)

preProcValues <- preProcess(BankTest, method = c("center", "scale"))

testTransformed <- predict(preProcValues, BankTest)
psych::describe(testTransformed)


```

## Fit Random Forest 
```{r,echo=TRUE}
set.seed(1)

BankRF<- train(
  form = factor(Bankrupt.) ~ .,
  data = BankTrain,
  #here we add classProbs because we want probs
  trControl = trainControl(method = "cv", number = 10,
                           classProbs =  TRUE),
  method = "rf",
  tuneLength = 3)#why 3?

BankRF
```


## Use the test set to make a prediction
```{r,echo=TRUE}
Bank_pred<-predict(BankRF,testTransformed)
head(BankRF)

```


<center>
# How can this be used?

### Random Forest does a great job predicting wether a comapny will go bankrupt or not in this scenario by using provided measurements. This prediction can help many different groups in regards to the longevity of future investments. Being able to make a accurate prediction of wether a organization is likely to file for bankruptcy could assist investors and help make more informed investments.

</center>



